ins---
title: "Milk Analysis"
author: "Gustavo Reyes"
date: "10/25/2021"
output: html_document
---
#Milk Analysis used for Pinto et al. 2023

```{r}
#Setting Working Directory-------------------------------------
setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) #Set working directory to the path of document
```

##Loading Libraries
```{r ,echo=FALSE}
#Opening Libary and Inputs-------------------------------------
source("Util_Library.R")
source("Functions_Full_Analysis.R")
```


##Loading Model Necessary Inputs. 
```{r}
library(tidyverse)
#Inputs and Source Files-------------------------------------

#Inputs
source("Main_Loop.R")
source("Input_Static.R")
source("Input_Functions.R")
source("Util_DFFunctions.R")
source("Util_DFWeekCreation.R")
source("Util_Counter&Lists.R")
#Functions
source("Util_Functions.R")
source("Util_CCFunctions2.R")
source("Util_VisualFunctions.R")
source("Util_Output_Functions.R")

```

## Generating Change from T and Temp Data
This chunk read the csv data for the milk over the 5 day period and the refrigerated control over the 5 day period
```{r}
#Milk 5 day period data
Time_Temp_Profile<-read.csv("5-Day-Milk-Data.csv")

#Milk Refrigerated control data
Time_Temp_Profile_RC<-read.csv("5-Day-Refrigerated-Control-Data.csv")
```

## Functions for growth and lag phases 
These functions are the ones that calculate the adjusted growth rate and adjusted lag phase based on the temperature at which the milk is exposed to. 
```{r}
#Function for growth and lag phase
new_growth_rate<-function(newTemp, oldMu,oldTemp = 6, T0 = -4.15){
  newMu<-((newTemp-T0)/(oldTemp-T0))* oldMu
  return (newMu)
}


#Calculation of the new lag time.
new_lag_time <- function (newTemp, oldLag, oldTemp = 6, T0 = -4.15) {
  numerator <- oldTemp -T0
  denom <- newTemp - T0
  newLag <- ( (numerator / denom)^2) * oldLag
  return(newLag)
}
```

## Functions to create a time and temperature profiles based on given parameters
This functions are not necessarily used in the model. However, they are here in case a different time and temperature profile for milk had to be created. 
```{r}
#Not in use, function to create a data frame based on interval and initial and final temperature
Time_Temp_Creation<-function(Total_Time, Interval, Initial_Temperature, Final_Temperature){
  Time_Temp_df<-data.frame("min"= seq(0,Total_Time,by = Interval),
                         "tempM"  = seq(Initial_Temperature,Final_Temperature, by = ((Final_Temperature - Initial_Temperature)/(length(seq(0,Total_Time,by = Interval)) - 1) )))
}

Time_Temp_Creation_Var<-function(Total_Time, Interval, Mean_Temperature, SD_Temperature){
  Time_Temp_df<-data.frame("min"= seq(0,Total_Time,by = Interval),
                         "tempM"= rnorm(n =Total_Time+1, mean = Mean_Temperature, sd= SD_Temperature))
  return (Time_Temp_df)
}
```

## Function to calculate growth of R100084 P Paoae using a buchanan 3 phase linear model.  
```{r}
#This function calculates thee growth based on a time and temperature profile for 1 specific milk with R100084 P Paoae
Func_Growth_LagCon<-function(In_Lag_Consumed,Time_Temp_df,Interval, AF){
  #In_Lag_Consumed= Total lag time consumed
  #Time_Temp_df = dataframe with time and temperature conditions
  #Interval = time interval in the time_temp_df in hrs. 
  Total_Lag_Consumed = In_Lag_Consumed
  Total_Growth = 0
  old_lag = 0
  NMax = 8.14 #P Paoae N max
  old_mumax = 0.083508 #P Paoae mu max at 6°C
  Growth_V = c()
  for (i in 1:nrow(Time_Temp_df)){
    if (Total_Lag_Consumed <1 && old_lag!=0){
      Lag_t_interval<-new_lag_time(newTemp = Time_Temp_df$tempM[i], oldLag = old_lag)
      Lag_Consumed<-Interval/Lag_t_interval
      Total_Lag_Consumed<-Total_Lag_Consumed+Lag_Consumed
      Growth = 0
    } else if (Total_Lag_Consumed>=1 | old_lag == 0){
      Growth = ((new_growth_rate(newTemp = Time_Temp_df$tempM[i], oldMu = old_mumax))/2.303)* AF #0.684 #Converted log10 from log ln
      Total_Growth = Total_Growth + (Growth*Interval)
    }
    Growth_V = c(Growth_V,Total_Growth)
    #print(length(Growth_V))
  }
  return(list(Total_Growth,Total_Lag_Consumed,Growth_V))
}

#Buchanan spoilage function
Spoilage_Function_Single_Milk<-function(Cont, Pop_Max, Time_Temp_df, Interval =1/60, AF){
  Lag_Consumed = 0
  #this function provides two outputs, the total growth, and the new updated lag phase consumed. 
  Output_Milk<-Func_Growth_LagCon(In_Lag_Consumed = Lag_Consumed ,Time_Temp_df = Time_Temp_df,Interval = Interval, AF=AF)
  Lag_Consumed = Output_Milk[[2]]
  Cont<-Output_Milk[[1]]+Cont
  if( Cont>Pop_Max){
     Cont = Pop_Max
  }
  return (list(Cont,Output_Milk[[3]]))
}
```


## Generating P Paoae population change as a factor of  the time and temperature conditions, considering the adjustment factor.

An adjustment factor of 1.32 was chosen to match the growth rate obtained to match the experimental trials for this study. 

Thus chunk generates change as a factor of time in system. 
```{r}
#Growth of milk over the ST 5 day condition
Output_Milk<-Spoilage_Function_Single_Milk(Cont = 2.44, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile, Interval =1/60,AF = 1.34)
Changes_Over_Time<-2.44+Output_Milk[[2]]
#plot(Changes_Over_Time)

#Growth of milk for the refrigerated control
Output_Milk_RC<-Spoilage_Function_Single_Milk(Cont = 2.44, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile_RC, Interval =1/60,AF = 1.34)
Changes_Over_Time_RC<-2.44+Output_Milk_RC[[2]]
plot(Changes_Over_Time_RC)

#calculating the milk with the fixed model Adjustment Factor 1.00. 

#Growth of milk over the ST 5 day condition
Output_Milk<-Spoilage_Function_Single_Milk(Cont = 2.44, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile, Interval =1/60,AF = 1)
Changes_Over_Time_NAF<-2.44+Output_Milk[[2]]
plot(Changes_Over_Time_NAF)

#Growth of milk using overnight storage
Output_Milk_RC<-Spoilage_Function_Single_Milk(Cont = 2.44, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile_RC, Interval =1/60,AF = 1)
Changes_Over_Time_RC_NAF<-2.44+Output_Milk_RC[[2]]
```

## Creates supplemental plot where we compare the  growth curves of the predicted plot with the adjustment factor vs. the experimental conditions

```{r}
df_rt_AF = data.frame("Share Table" = Changes_Over_Time,
                      "Refrigerated Control" = Changes_Over_Time_RC[1:length(Changes_Over_Time)],
                      "min" = 1:length(Changes_Over_Time))

df_rt_AF_melted<-melt(df_rt_AF,id.vars = "min")
df_rt_AF_melted$AF = "Adjusted (Factor = 1.34)"

df_rt_NAF = data.frame("Share Table" =Changes_Over_Time_NAF,
                      "Refrigerated Control" = Changes_Over_Time_RC_NAF[1:length(Changes_Over_Time_NAF)],
                      "min" = 1:length(Changes_Over_Time_NAF))

df_rt_NAF_melted<-melt(df_rt_NAF,id.vars = "min" )
df_rt_NAF_melted$AF = "Not Adjusted"

December_Growth_Data_R1<-data.frame("min" = c(0,125,1440,1565,2880,3005,4320,4195,5760,5885), 
                                    "counts" = c(2.53,2.48, 3.65,3.95,4.87, 5.18, 6.3,6.32,6.89,7.1))

December_Growth_Data_R2<-data.frame("min" = c(0,125,1440,1565,2880,3005,4320,4195,5760,5885), 
                                    "counts" = c(2.51,2.81, 3.69, 3.76, 5.02, 5.11, 6.24,6.41,7.19,7.15))

December_Growth_Data_R3<-data.frame("min" = c(0,125,1440,1565,2880,3005,4320,4195,5760,5885), 
                                    "counts" = c(2.62, 2.9,3.62,3.59,  4.94, 5.12,6.16, 6.38,7.13,7.2))

December_Growth_Data_Comb = rbind(December_Growth_Data_R1,December_Growth_Data_R2,December_Growth_Data_R3)


December_RC_Data_Comb<-data.frame("min" = c(0,1440,2880,4320,5760), 
                                    "counts" = c(2.44, 3.25,4.31,5.39,6.19))

#For  Both winter and summer trials
All_Seasons_RC_Data_Comb<-data.frame("min" = c(0,0,0,
                                               1440,1440,1440,
                                               2880,2880,2880,
                                               4320,4320,4320,
                                               5760,5760,5760), 
                                    "counts" = c(2.44,2.31,2.31,
                                                 2.47,2.7,3.25,
                                                 3.31,3.64,4.31,
                                                 4.47,4.48,5.39,
                                                 5.31,5.31,6.19))



Data_For_Plot = rbind(df_rt_AF_melted,df_rt_NAF_melted)

ggplot(aes(x = min, y = value), data = Data_For_Plot)+
geom_line(aes(color = variable), linewidth = 1)+
facet_wrap(~AF)+
theme_bw()+
theme(legend.position="bottom",legend.box="vertical")+
geom_point(data = December_Growth_Data_Comb, aes(x  = min, y = counts, fill = "Experimental ST System 2hr"), shape = 21)+
geom_point(data = December_RC_Data_Comb, aes(x  = min, y = counts,fill = "Experiment Refrigerated Control"), shape = 22)+
scale_fill_manual(name = "",values = c("Experimental ST System 2hr" = "coral", "Experiment Refrigerated Control" = "darkblue"))+
scale_color_discrete(name = "",labels=c('Modeled ST System 2hr',"Experiment Refrigerated Control"))+
  geom_hline(yintercept = 6.0, color = "black", linewidth = 1)+
geom_text(aes(2000,5.8,label = "Quality Threshold = 6.0"), color = "black", size  = 3,check_overlap = TRUE)+
  labs(x= "Residence time in system (min)", y = expression("Average log"[10]~"CFU/ml"))

ggsave("Figures/Supplemental-Predicted vs Experimental.jpg", width =7, height = 5, units = "in", dpi = 300)

```


##Running the  Model for Milk.
```{r warning=TRUE}
#Running Loop First:-------------------------------------
start_time<-Sys.time()
source("Main_Loops2.R")
```

## This converts the list of dataframes from the simulation into a full clean dataframe, 
```{r}
#1. Start from here
Individual_Analysis<-rbind.fill(List_Sens_Pre)

#2. find the duplicates
#this step filters replicated based on the ID
Individual_Analysis<-Individual_Analysis %>% 
  group_by(ID) %>% 
  filter(TotServices==max(TotServices))
#3
AnalysysDF<-Individual_Analysis
AnalysisDF_Summer<-Individual_Analysis
AnalysysDF_2<-Individual_Analysis

#Adding the time of the services for those milks that were added to the service line during the lunch period. The residence time is ajusted to show all milks were set outusde refrigeration for the whole lunch period. 
AnalysysDF[AnalysysDF$Initial.Service>1,]
AnalysysDF$TotTime<-AnalysysDF$TotTime+(as.integer(AnalysysDF$Initial.Service)-1)*(Time_Service_Length + Time_Turnaround_Length)
AnalysysDF$Shared<-ifelse(AnalysysDF$STtimes>0, "Yes", "No")


#Updating the summer one
#Adding the time of the services for those milks that were added to the service line during the lunch period. The residence time is ajusted to show all milks were set outusde refrigeration for the whole lunch period. 
AnalysisDF_Summer[AnalysisDF_Summer$Initial.Service>1,]
AnalysisDF_Summer$TotTime<-AnalysisDF_Summer$TotTime+(as.integer(AnalysisDF_Summer$Initial.Service)-1)*(Time_Service_Length + Time_Turnaround_Length)
AnalysisDF_Summer$Shared<-ifelse(AnalysisDF_Summer$STtimes>0, "Yes", "No")
```


## Updating Final levels in milk based on the growth observed as the total residence time in milk
```{r}
#Winter
Output_Milk<-Spoilage_Function_Single_Milk(Cont = 0, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile, Interval =1/60,AF = 1.32)
Changes_Over_Time<-Output_Milk[[2]]
#max(Changes_Over_Time)

NMax = 8.14
for (i in 1:nrow(AnalysysDF)){
  In_Cont<-AnalysysDF$InSpoilageCon[i]
  Time<-AnalysysDF$TotTime[i]
  if (Time == 0){
    Total_Growth = 0
  }else{
    Total_Growth<-Changes_Over_Time[Time]
  }
  Final_Con<-In_Cont+Total_Growth
  if(Final_Con>NMax){
    Final_Con <- NMax
  }
  AnalysysDF$SpoilageCon[i]<-Final_Con
}

#Summmer..
Output_Milk_Summer<-Spoilage_Function_Single_Milk(Cont = 0, Pop_Max =8.14, Time_Temp_df = Time_Temp_Profile, Interval =1/60,AF = 1.32)
Changes_Over_Time_Summer<-Output_Milk_Summer[[2]]

```

## Analysis and Generating outputs

#Relationship between Time and APC Levels. Main Figure
```{r}
p_1_log<-AnalysysDF%>% 
  ggplot(aes(y = SpoilageCon, x= TotTime, fill= Shared, shape = Shared))+
  geom_point( size =2, alpha = 0.7, color = "black")+
  geom_hline(yintercept =6)+
geom_text(aes(100,6.5,label = "Quality Threshold = 6.0"), color = "black", size  = 3,check_overlap = TRUE)+
  facet_wrap(.~Location)+
  labs(x = "Residence Time in Cafeteria System (min)", y = expression("Pseudomonas poae population (log"[10]~"CFU/ml)"))+
  scale_y_continuous()+
  scale_shape_manual(values = c(21,23))+
  scale_fill_manual(values = c("coral1", "seagreen3"))+
  scale_x_log10(n.breaks= 10)+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = -45, vjust = -0.3))+
  theme(legend.position="bottom")+
  geom_segment(aes(x = 1440 , y = -Inf, xend = 1440, yend = 6,linetype = "Start of Day"))+
  geom_segment(aes(x = 2880 , y = -Inf, xend = 2880, yend = 6), linetype = "dotted") +
  geom_segment(aes(x = 4320 , y = -Inf, xend = 4320, yend = 6), linetype = "dotted") +
  geom_segment(aes(x = 0 , y = -Inf, xend = 0, yend = 6), linetype = "dotted")  +
  geom_segment(aes(x = 5760 , y = -Inf, xend = 5760, yend = 6), linetype = "dotted") +
  scale_linetype_manual(name = "",values = c("dotted"))

#p_1_log

#ggsave("Figures/Share Table Spoilage-log.jpg", width =8, height = 5, units = "in", dpi = 300)


p_1<-AnalysysDF%>%
  ggplot(aes(y = SpoilageCon, x= TotTime, fill= Shared, shape = Shared))+
  geom_point( size =2, alpha = 0.7, color = "black")+
  geom_hline(yintercept =6)+
geom_text(aes(2000,6.5,label = "Quality Threshold = 6.0"), color = "black", size  = 3,check_overlap = TRUE)+
  facet_wrap(.~Location)+
  labs(x = "Residence Time in Cafeteria System (min)", y = expression("Pseudomonas poae population (log"[10]~"CFU/ml)"))+
  scale_y_continuous()+
  scale_shape_manual(values = c(21,23))+
  scale_fill_manual(values = c("coral1", "seagreen3"))+
  #scale_x_log10(n.breaks= 12)+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = -45, vjust = -0.3))+
  theme(legend.position="bottom")+
  geom_segment(aes(x = 1440 , y = -Inf, xend = 1440, yend = 6,linetype = "Start of Day") )+
  geom_segment(aes(x = 2880 , y = -Inf, xend = 2880, yend = 6), linetype = "dotted") +
  geom_segment(aes(x = 4320 , y = -Inf, xend = 4320, yend = 6), linetype = "dotted") +
  geom_segment(aes(x = 0 , y = -Inf, xend = 0, yend = 6), linetype = "dotted") +
  geom_segment(aes(x = 5760 , y = -Inf, xend = 5760, yend = 6), linetype = "dotted") +
  scale_linetype_manual(name = "",values = c("dotted"))

#p_1

#ggsave("Figures/Share Table Spoilage-NonLog.jpg", width =8, height = 5, units = "in", dpi = 300)
```

## Code to Generate Summary Tables S3 aand S4.

# Table S3 and Table S4
```{r}

#Table s3 and Table S4

#Overall
Overall_s3<-AnalysysDF %>% 
  #filter(Shared == "No") %>% 
  summarize(TimeMedian= median(TotTime), 
            Timeq25 = quantile(TotTime, 0.025), 
            Timeq975 = quantile(TotTime, 0.975),
            PoPChangeMedian = median(SpoilageCon - InSpoilageCon),
            PopChangeq25 =  quantile(SpoilageCon - InSpoilageCon, 0.025),
            PopChange1975 = quantile(SpoilageCon - InSpoilageCon, 0.975),
            TotalSpoiled = sum(SpoilageCon>6),
            Per_Spoiled = (sum(SpoilageCon>6)/nrow(.))*100
            ) %>% 
  mutate(Type = "Overall")


NotShared_s3<-AnalysysDF %>% 
  filter(Shared == "No") %>% 
  summarize(TimeMedian= median(TotTime), 
            Timeq25 = quantile(TotTime, 0.025), 
            Timeq975 = quantile(TotTime, 0.975),
            PoPChangeMedian = median(SpoilageCon - InSpoilageCon),
            PopChangeq25 =  quantile(SpoilageCon - InSpoilageCon, 0.025),
            PopChange1975 = quantile(SpoilageCon - InSpoilageCon, 0.975),
            TotalSpoiled = sum(SpoilageCon>6),
            Per_Spoiled = (sum(SpoilageCon>6)/nrow(.))*100
            ) %>% 
  mutate(Type = "Not Shared")

Shared_s3<-AnalysysDF %>% 
  filter(Shared == "Yes") %>% 
  summarize(TimeMedian= median(TotTime), 
            Timeq25 = quantile(TotTime, 0.025), 
            Timeq975 = quantile(TotTime, 0.975),
            PoPChangeMedian = median(SpoilageCon - InSpoilageCon),
            PopChangeq25 =  quantile(SpoilageCon - InSpoilageCon, 0.025),
            PopChange1975 = quantile(SpoilageCon - InSpoilageCon, 0.975),
            TotalSpoiled = sum(SpoilageCon>6),
            Per_Spoiled = (sum(SpoilageCon>6)/nrow(.))*100
            ) %>% 
  mutate(Type = "Shared")


Table_s3<-rbind(Overall_s3, NotShared_s3, Shared_s3)
Table_s3


### Table S4
AnalysysDF %>% 
  group_by(Location) %>% 
  summarise(count = n(),
            percentage = count/nrow(.),
            day1= sum(TotTime<1440),
            day1_per = sum(TotTime<1440)/nrow(.),
            day2 =sum(TotTime>1440 & TotTime<1440*2),
            day2_per = sum(TotTime>1440 & TotTime<1440*2)/nrow(.),
            day3= sum(TotTime>1440*2 & TotTime<1440*3),
            day3_per = sum(TotTime>1440*2 & TotTime<1440*3)/nrow(.),
            day4= sum(TotTime>1440*3 & TotTime<1440*4),
            day4_per = sum(TotTime>1440*3 & TotTime<1440*4)/nrow(.),
            day5= sum(TotTime>1440*4 & TotTime<1440*5),
            day5_per = sum(TotTime>1440*4 & TotTime<1440*5)/nrow(.),
            )


#Summary for paper line 448
AnalysysDF %>% 
  filter(Shared == "Yes") %>% 
  group_by(Location) %>% 
  summarise(count = n(),
            per = (count/nrow(.))*100)

AnalysysDF %>% 
  #filter(Shared == "Yes") %>% 
  group_by(Location) %>% 
  summarise(count = n(),
            per = (count/nrow(.))*100)

```




# ADDITIONAL ANALYSES


###How many Items ended up in which locations
```{r}
library(scales)
library(forcats)

AnalysysDF %>% 
  count(Location) %>% 
  arrange(desc(n)) %>% 
  mutate(Location = fct_reorder(Location, n)) %>%
  ggplot(aes(x =Location, y = n/sum(n)*100))+
  scale_y_continuous(labels = scales::percent_format(scale = 1) )+
  geom_col()+
  labs(x = "Final Food Location", y = "Percent of milk items by location", title= "Final Location for milk-Donation happens end of week")+
  theme(plot.title = element_text(hjust = 0.5))+
  coord_flip()
  
```

###What is the distribution of time for the milk items

```{r}
AnalysysDF %>%
  ggplot(aes(x = TotTime))+
  geom_histogram()+
  scale_x_continuous(n.breaks= 10)+
  labs(x = "Time in Minutes", y = "Count ", title= "Time in system by Item week")+
  theme(plot.title = element_text(hjust = 0.5))
  

```

###APC Contamination by Items

```{r}
AnalysysDF %>%
  ggplot()+
  geom_histogram(aes(x = SpoilageCon),color="darkblue", fill="lightblue",alpha= 0.3)+
  geom_histogram(aes(x = InSpoilageCon),color="darkblue", fill="lightgreen", alpha= 0.3)+
  geom_vline(xintercept =6, color= "Red")+
  scale_x_continuous(n.breaks= 10)+
  labs(x = "Phsychotroph Population Log CFU/ml", y = "Count", title= "Levels per milk carton")+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
AnalysysDF %>% 
  group_by(ConsumedAt) %>% 
  summarise(ChangeCon = SpoilageCon-InSpoilageCon) %>% 
  filter(ConsumedAt!="") %>% 
  ggplot(aes(x = ChangeCon, fill = ConsumedAt))+
  geom_histogram(bins=50)+
  labs(x = "Change in population log CFU/ml")

```




#Change in Contamination Levels due to Time
  ##Need to figure out a way to track contamination change. 
```{r}
AnalysysDF %>% 
  mutate(Change_Cont = SpoilageCon-InSpoilageCon) %>% 
  ggplot(aes(y = Change_Cont, x= TotTime, fill = Shared))+
  geom_point( size =2, shape = 21, alpha = 0.7)+
  facet_wrap(.~Location)+
  labs(x = "Residence Time in Cafeteria System (min)", y = expression("Pseudomonas poae population Change (log"[10]~"CFU/ml"))+
  scale_y_continuous()+
  scale_x_log10(n.breaks= 10)+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_bw()+
  theme(legend.position="bottom")+
  theme(axis.text.x = element_text(angle = -45, vjust = -0.3))

#ggsave("growth over time 2.0.jpg", width =8, height = 5, units = "in", dpi = 300)


```
