---
title: "Summer Spoilage Study"
author: "Gabriella Pinto"
date: "2023-08-15"
output: html_document
---

## Load libraries for analysis and plots

```{r}
library(readr)
library(dplyr)
library(grid)
library(ggh4x)
library(ggplot2)
library(ggpattern)
library(ggpubr)
library(gmodels)
library(Hmisc)
library(tidyr)
```

## Import Data & Initial Data Cleaning for Figure 1: Microbial Counts

```{r}
log_szns <- read_csv("C:/Users/Gabby/Box/Share Table Project/Milk Spoilage Model Docs/Spoilage Study December/log_szns.csv")

results_both = filter(log_szns)%>%
  filter(!is.na(log_szns$sma))%>%
  arrange(desc(tod))
  
results_both$rep = as.factor(results_both$rep)

results_both$tod[results_both$tod == "Start"] = "Start of First Service"
results_both$tod[results_both$tod == "End"] = "End of Second Service"
```

## Take data out of log scale for calculations

```{r}
results_both$raw = 10^(results_both$sma)
```

## Calculation for Error Bars

#Writing log function with rounding
```{r}
baselog = function(x){
  round(log10(x),2)
}
```

#Split into inoculated data frame, calculate mean, min and max for error bars

```{r}
ino_errors = filter(results_both)%>%
  filter(type == "Inoculated A" | type == "Inoculated B" | type == "Inoculated"| type == "Refrigerated Control Inoculated" | type == "Refrigerated Control A" | type == "Refrigerated Control B")%>%
  group_by(day, tod, Season,type)%>%
  summarise (mean = baselog(mean(raw)), min = baselog(min(raw)), max = baselog(max(raw)))
```

#Perform necessary data cleaning for plot
```{r}
ino_errors$Inoculated = "Inoculated Milk"
ino_errors$type[ino_errors$type == "Inoculated A"] = "Inoculated"
ino_errors$type[ino_errors$type == "Inoculated B"] = "Inoculated"
ino_errors$type[ino_errors$type == "Refrigerated Control Inoculated"] = "Refrigerated Control"
ino_errors$type[ino_errors$type == "Refrigerated Control A"] = "Refrigerated Control"
ino_errors$type[ino_errors$type == "Refrigerated Control B"] = "Refrigerated Control"
ino_errors$tod<-factor(ino_errors$tod, levels = c("Start of First Service", "End of Second Service"))
```

#Split into NOT inoculated data frame, calculate mean, min and max for error bars

```{r}
noino_errors = filter(results_both)%>%
  filter(type == "Not Inoculated" | type == "Refrigerated Control Not Inoculated")%>%
  group_by(day, tod, Season,type)%>%
  summarise (mean = baselog(mean(raw)), min = baselog(min(raw)), max = baselog(max(raw)))
```

#Perform necessary data cleaning for plot
```{r}
noino_errors$`Not Inoculated` = "Uninoculated Milk"
noino_errors$type[noino_errors$type == "Refrigerated Control Not Inoculated"] = "Refrigerated Control"
noino_errors$tod<-factor(noino_errors$tod, levels = c("Start of First Service", "End of Second Service"))
```

## Generate plots for Figure 1

#Inoculation Figure 1

```{r}
sma_ino_eb = ino_errors %>%
  ggplot(aes(x =`day`, y = `mean`, color = `type`, shape = `tod`))+
  geom_point(position = position_dodge(width =0.7), size = 2, stroke = 1.5)+
  geom_errorbar(aes(ymin =`min`, ymax=`max`,color = `type`, shape = `tod`),position = position_dodge(width =0.7), size = 0.5)+
  xlab("Day")+
  xlim(0.5,5.5)+
  ylim(2,7.5)+
  ylab((expression("Counts log"[10]~"(CFU/ml)")))+
  scale_color_discrete(name = "Sample Type")+
  scale_shape_manual(name = "Collection Time", values = c(16, 2), guide = guide_legend(nrow = 1, ncol =2, byrow =TRUE))+
  theme_bw()+
  facet_grid2(Inoculated~Season, switch = "y")+
  theme(legend.title= element_text(size=9,face="bold"), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(face="bold", hjust =0.5), 
        legend.text = element_text(size=9),
        legend.position = "top",
        legend.key = element_rect(fill="whitesmoke"), 
        legend.background = element_rect(fill=NA),
        #legend.direction = "vertical",
        legend.margin = margin())+
  geom_hline(yintercept = 6, color = "black", linewidth = 0.8)+
  geom_text(aes(3,5.5,label = "Spoilage Threshold = 6.0", vjust = -1), color = "black", size  = 3, position = position_dodge(width =1), check_overlap = TRUE)+
  geom_hline(yintercept = 2.31, color = "black", linewidth =1, linetype = "dashed")+
  geom_text(aes(3,2.31,label = "LOD = 2.31", vjust = -1), color = "black", size  = 3,position = position_dodge(width = 1), check_overlap = TRUE)

sma_ino_eb
```


#No inoculation Figure 1

```{r}
sma_noino_eb = noino_errors %>%
  ggplot(aes(x =`day`, y = `mean`, color = `type`, shape = `tod`))+
  geom_point(position = position_dodge(width =0.7), size = 2, stroke = 1.5)+
  geom_errorbar(aes(ymin =`min`, ymax=`max`,color = `type`, shape = `tod`),position = position_dodge(width =0.7), size = 0.5)+
  xlab("Day")+
  xlim(0.5,5.5)+
  ylim(2,7.5)+
  ylab((expression("Counts log"[10]~"(CFU/ml)")))+
  scale_color_discrete(name = "Sample Type")+
  scale_shape_manual(name = "Collection Time", values = c(16, 2), guide = guide_legend(nrow = 3, ncol =1, byrow =TRUE))+
  theme_bw()+
  facet_grid2(`Not Inoculated`~Season, switch = "y")+
  theme(legend.title= element_text(size=9,face="bold"), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(face="bold", hjust =0.5), 
        legend.text = element_text(size=9),
        legend.position = "right",
        legend.key = element_rect(fill="whitesmoke"), 
        legend.background = element_rect(fill=NA),
        legend.direction = "vertical",
        legend.margin = margin())+
  geom_hline(yintercept = 6, color = "black", linewidth = 0.8)+
  geom_text(aes(3,5.5,label = "Spoilage Threshold = 6.0", vjust = -1), color = "black", size  = 3, position = position_dodge(width =1), check_overlap = TRUE)+
  geom_hline(yintercept = 2.31, color = "black", linewidth =1, linetype = "dashed")+
  geom_text(aes(3,2.31,label = "LOD = 2.31", vjust = -1), color = "black", size  = 3,position = position_dodge(width = 1), check_overlap = TRUE)

sma_noino_eb
```

#Export Figure 1
```{r}
require(grid)
szn_plot_eb = ggarrange(sma_ino_eb + rremove("ylab") + rremove("xlab"), sma_noino_eb + rremove("ylab") + rremove("xlab"),
                           ncol = 1, nrow =2,
                           legend.grob = get_legend(sma_ino_eb),
                           legend = "top",
                           widths = 10,
                           heights = 20)

  annotate_figure(szn_plot_eb, 
                  left = textGrob((expression("Counts log"[10]~"(CFU/ml)")), 
                                  rot = 90, 
                                  vjust = 1),
                                 
                  bottom = textGrob("Day"))%>%
    ggexport(filename = "Figure 1.tiff", res = 250, width = 2500, height = 2000, compression = "lzw")

```


## Import Data & Initial Data Cleaning for Figure 2: Log Shifts


```{r}
szn_shift <- read_csv("C:/Users/Gabby/Box/Share Table Project/Milk Spoilage Model Docs/Spoilage Study December/szns_shift.csv")

shiftnew2 = szn_shift

shiftnew2$`shift type`[shiftnew2$`shift type` == "overnight"] = "Overnight"
shiftnew2$`shift type`[shiftnew2$`shift type` == "share table"] = "Share Table"

shiftnew2$day = as.factor(shiftnew2$day)
shiftnew2$Inoculated = "Inoculated Milk"
shiftnew2$`Not Inoculated` = "Uninoculated Milk"

```

# Creating new data frame with all data points to overlay for inoculated plot

```{r}
points2 = shiftnew2 %>%
  filter(media == "sma" & type == "Inoculated")

```

# Creating new data frame with all data points to overlay for NOT inoculated plot
```{r}
points3 = shiftnew2 %>%
  filter(media == "sma" & type == "Not Inoculated")
```


## Generate plots for Figure 2

```{r}
hist_both = shiftnew2 %>% 
  filter(type == "Inoculated" & media == "sma") %>%
  ggplot(aes( `day`, `log shift`, color = `shift type` ))+
    facet_grid2(Inoculated~`Season`, switch = "y")+
    geom_boxplot(varwidth = TRUE)+
    scale_color_discrete(name = "Storage Type", guide = guide_legend(reverse = TRUE, nrow =1 , ncol = 3, byrow =TRUE))+
    geom_point(data = points2, aes(x= `day`, y = `log shift`, color = `shift type`), position = position_dodge(width =0.5), size = 1.25)+
    #geom_jitter(shape = 16, size = 2, position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.5))+
    theme_bw()+
    theme(legend.title= element_text(size=9,face="bold"), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(face="bold"), 
        legend.text = element_text(size=9),
        legend.position = "bottom",
        legend.key = element_rect(fill="whitesmoke"), 
        legend.background = element_rect(fill=NA),
        #legend.direction = "vertical",
        legend.margin = margin())+
    #guides(color=guide_legend(nrow=1, ))+
    xlab("Day")+
    ylab((expression("Change in Microbial Counts log"[10]~"(CFU/ml)")))

hist_both
```

```{r}
hist_both_no = shiftnew2 %>% 
  filter(type == "Not Inoculated" & media == "sma") %>%
  ggplot(aes( `day`, `log shift`, color = `shift type` ))+
    facet_grid2(`Not Inoculated`~`Season`, switch = "y")+
    geom_boxplot(varwidth = TRUE)+
    scale_color_discrete(name = "Storage Type", guide = guide_legend(reverse = TRUE, nrow =1, ncol =3, byrow =TRUE))+
    geom_point(data = points3, aes(x= `day`, y = `log shift`, color = `shift type`), position = position_dodge(width =0.5), size = 1.25)+
    #geom_jitter(shape = 16, size = 2, position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.5))+
    theme_bw()+
    theme(legend.title= element_text(size=9,face="bold"), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(face="bold"), 
        legend.text = element_text(size=9),
        legend.position = "bottom",
        legend.key = element_rect(fill="whitesmoke"), 
        legend.background = element_rect(fill=NA),
        #legend.direction = "vertical",
        legend.margin = margin())+
    #guides(color=guide_legend(nrow=3))+
    xlab("Day")+
    ylab((expression("Change in Microbial Counts log"[10]~"(CFU/ml)")))

hist_both_no

```
```{r}
ggsave("C:/Users/Gabby/Downloads/logshift_hist_no_both.tiff", plot = hist_both_no, width = 7.5, height = 5, units = "in", dpi = 1000)
```

#Export Figure 2

```{r}
szn_shift_plot = ggarrange(hist_both + rremove("ylab") + rremove("xlab"), hist_both_no + rremove("ylab") + rremove("xlab"),
                           ncol = 1, nrow =2,
                           legend.grob = get_legend(hist_both),
                           legend = "top",
                           widths = 10,
                           heights = 20)

  annotate_figure(szn_shift_plot, 
                  left = textGrob((expression("Change in Microbial Counts log"[10]~"(CFU/ml)")), 
                                  rot = 90, 
                                  vjust = 1),
                                 
                  bottom = textGrob("Day"))%>%
    ggexport(filename = "Figure 2.tiff", res = 250, width = 2500, height = 2000, compression = "lzw")
```

